### 2021-10-11 docs/paper/sections/3_approach.tex
Improve black/white accessibility of figure 2
Minor rephrasings and fixed throughall the paper

\begin{figure*}
	\begin{minipage}{\linewidth}
		\newcommand\textlowlight[1]{\textcolor{gray}{#1}}
		\newcommand\texthighlight[1]{\textcolor{accent1}{#1}}
		\begin{center}
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/CallExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/PropertyAccessExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/JsxOpeningElement}
			\end{subfigure}

@@

To do so, we start by parsing the source code of every dependency repository as well as the source code of the target package each into a separate abstract syntax forest (ASF, i.e., a set of ASTs).

In a second step, we perform a static type analysis against each dependency ASF together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node will contain a link to the assignment node of this variable, and every function call expression will contain a link to the definition of this function.
The operating principle of this type analysis depends on the kind of programming language; i.e., in statically typed languages, type symbols can usually be retrieved from the declaration of an identifier, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

>>>>

\begin{figure*}
	\begin{minipage}{\linewidth}
		\colorlet{lowlight}{black!40}
		\newcommand\textlowlight[1]{\textcolor{lowlight}{#1}}
		\newcommand\texthighlight[1]{\textbf{\textcolor{accent1}{#1}}}
		\begin{center}
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/CallExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/PropertyAccessExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/JsxOpeningElement}
			\end{subfigure}

@@

To do so, we start by parsing the source code of every dependency repository as well as the source code of the target package each into a separate \emph{abstract syntax forest}F (ASF, i.e., a set of ASTs).

In a second step, we perform a static type analysis against each dependency ASF together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node will contain a link to the assignment node of this variable, and every function call expression will contain a link to the definition of this function.
The operating principle of this type analysis depends on the kind of programming language; that is, in statically typed languages, type symbols can usually be retrieved from the declaration of an identifier, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

### 2021-10-11 docs/paper/sections/4_implementation.tex
Minor rephrasings and fixed throughall the paper

The de facto standard package repository for node.js is npm.
To fetch downstream dependency packages from the npm registry, we use the npm package \code{npm-dependants}\footnote{All packages can be retrieved from \url{npmjs.com}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, including the \emph{tarball} URL of the latest package version, using the package \code{package-metadata}.
Finally, we download and extract each tarball using the package \code{download-package-tarball}.

@@

For the second approach based on a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a certain package, we can search all available \code{package.json} files for a reference to this package identifier.
We refine our search query with some additional constraints to filter out package copies from \code{node_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved.
We submit this search query to the publicly available GraphQL\footnote{\url{https://graphql.com/}} API of Sourcegraph\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package \code{graphql-request}.
After retrieving a search result, we download a snapshot of the full repository by using the package \code{download-git-repo}.

@@

@@

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to the \emph{Visual Studio Code} IDE\footnote{\url{https://code.visualstudio.com/}}.
VS Code is one of the most popular IDEs that support JavaScript, and the one with the highest annual growth.\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
It also provides a comprehensive set of APIs for extension developers.
To support other researchers in reusing our solution, we also provide a CLI.

To support developers in answering the questions raised in \cref{sec:approach}, we have designed three key views for our extension (see \cref{fig:implementation/presentation/screenshots}):

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm feeling lucky'' button for every view that redirects the user to a random dependency or reference, resp., to gain a faster, unbiased impression of usage samples.

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the \emph{Tree View API} and the \emph{CodeLens API}.
One challenge has been to deliver fast results and not to block the user interface because even in spite of the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation}).
To overcome this issue, we use pagination wherever possible and apply the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our object model by using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.

>>>>

The de facto standard package repository for node.js is npm.
To fetch downstream dependency packages from the npm registry, we use the npm package \code{npm-dependants}\footnote{All packages can be retrieved from \url{npmjs.com}.} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, including the \emph{tarball} URL of the latest package version, using the package \code{package-metadata}.
Finally, we download and extract each tarball using the package \code{download-package-tarball}.

@@

For the second method based on a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a \code{package.json} metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a certain package, we can search all available \code{package.json} files for a reference to this package identifier.
We refine our search query with some additional constraints to filter out package copies from \code{node_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved.
We submit this search query to the publicly available GraphQL\footnote{\url{https://graphql.com/}} API of Sourcegraph\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package \code{graphql-request}.
After retrieving a search result, we download a snapshot of the full repository by using the package \code{download-git-repo}.

@@

After all references have been collected, a proper user interface (UI) is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to the \emph{Visual Studio Code} IDE\footnote{\url{https://code.visualstudio.com/}}.
VS Code is one of the most popular IDEs that support JavaScript, and the one with the highest annual growth.\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
It also provides a comprehensive set of APIs for extension developers.
To support other researchers in reusing our solution, we also provide a CLI.

To support developers in answering the questions raised in \cref{sec:approach}, we have designed three key views (see \cref{fig:implementation/presentation/screenshots}):

@@

Both references and members are organized each in a tree view that reflects the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm feeling lucky'' button for every view that redirects the user to a random dependency or reference, resp., to gain a faster, unbiased impression of usage samples.

To implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the \emph{Tree View API} and the \emph{CodeLens API}.
One challenge has been to deliver fast results and not to block the user interface because even in spite of the lightweight mining methods we chose to satisfy \cref{req1} from above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation}).
To overcome this issue, we use pagination wherever possible and apply the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our object model by using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting during the data collection process.

### 2021-10-11 docs/paper/sections/5_evaluation.tex
Minor rephrasings and fixed throughall the paper

\begin{enumerate}[label=RQ\arabic*]
	\item \label[requirement]{resqu1} What is the quality and quantity of the proposed approaches for dependency collection?
	\item \label[requirement]{resqu2} What is the quality and quantity of the proposed approach for mining usage samples?
	\item \label[requirement]{resqu3} How well is the proposed tool applicable?
\end{enumerate}

@@

To assess the quality of dependency collection, we analyze it with respect to extent, precision, recall, and performance.
\Cref{tab:evaluation/resqu1/quantities} shows the number of dependencies that we collected to build an experimental dataset from each data source following the proposed approaches for a set of arbitrarily selected npm packages.
For each package, we have annotated a subset of the collected dependencies (max. 20 dependencies per package) to identify and classify false positive hits.

\begin{table}
	\caption{Quantity and false-positive rates (FPR) of downstream dependencies found by both approaches for selected packages.}
	\label{tab:evaluation/resqu1/quantities}

@@

Regarding the total number of dependencies collected by each approach, no clear trend in favor of any approach can be ascertained for small or medium packages (having less than \num{10,000} stars on GitHub).
For larger packages (having at least \num{10,000} stars), however, the number of dependencies collected from the npm registry stagnates near to 400 hits; beyond this limit, the npm registry will return internal server errors only.
For Sourcegraph, we have not hit any limitations so far; currently, they do not provide any official documentation for the exact rate limits.

As the disjunct proportion of results from both approaches is very small (on average about \SI{6}{\percent}), we consider a combination of both data sources useful for maximizing the extent and diversity of the gained dataset.

@@

In some situations, this may happen if the package is a plugin for another package or if it is invoked as a CLI from a build script, but in the majority of repositories, developers trivially appear to have specified the upstream dependency by accident (e.g., while copying a \code{package.json} file over from another package), or to have forgotten to remove the upstream dependency after switching away from using the package.
Rarer causes include repositories that only add type definitions for TypeScript to a package but do not actually import it, as well as, especially for repositories found on Sourcegraph, other packages required by a repository declaring a peer dependency on the target package which needs to be fulfilled by the depending repository.

@@

Besides the precision of collected dependencies, their recall is also of interest.
As the dependencies are collected from two giant datasets in excerpts only, a quantitative analysis of false-negative hits based on manual annotation of these datasets would be too expensive for this paper.
Nevertheless, there are some causes that will prevent a dependency from being found by our approaches:

\begin{itemize}
	\item Packages without a proper dependency manifest cannot be detected by our approaches which are based on exactly these metadata.
	\item On npm, only published packages will be found, creating a bias towards generic and professional software.
	\item On both platforms, results will be sorted based on intransparent criteria (which, as we speculate, include the number of direct dependents on npm and the recent update frequency on Sourcegraph, resp.).
		As we only fetch the first dependencies from both data sources, this is a likely source of further biases.
		These could be fought by always fetching all dependencies, but this would drastically reduce the overall performance.
\end{itemize}

@@

To assess the quality of usage mining, we examine the precision, recall, and performance of this processing step.
Due to the complexity of a detailed annotation process, we only consider the existence of found usage samples on the dependency level.
The absolute quantity of found usage samples varies largely for different packages and dependencies based on the semantic extent of the package and the coupling between package and dependency.

@@

Under laboratory conditions, false-positive usage samples cannot be emitted by our approach if we assume the correctness of the TypeScript compiler.
There are only two exceptions to this invariant:
\begin{enumerate*}[label=(\roman*)]
	\item In case of a name collision between two packages, invalid or false positives usage samples may be emitted.
		However, the npm infrastructure attempts to rule out these collisions by using the package name as a unique ID for every published package.
	\item If dependency developers deliberately interfere with the TypeScript compiler (for instance, by suppressing type errors using a \verb|@ts-ignore| comment\footnote{\url{https://www.typescriptlang.org/docs/handbook/release-notes/typescript-2-6.html\#suppress-errors-in-ts-files-using--ts-ignore-comments}}), both false positive and false negative matches are possible.
\end{enumerate*}

@@

To evaluate the recall of our approach, we have refined the experimental dataset collected in \cref{sec:evaluation/resqu1} and removed all false-positive dependencies.
Within this cleansed dataset, the overall rate of dependencies for that our approach is unable to find at least one usage sample accounts for \SI{47.9}{\percent}.
As a cause for these false negatives, we have identified a number of systematic causes:

\begin{itemize}
	\item Many projects on OSS platforms have complex build configurations that include additional transpilation or code generation steps before a final version of the source code is reached that is valid to the node.js interpreter or the TypeScript compiler.
		Unless we add explicit support for such build configurations, type analysis and usage mining for these projects will fail.
	\item In some situations, additional type definitions are required to perform a complete type inference of a dependency.
		This applies to every parametrized callback from a package for that type definitions are not available.
		This limitation could be resolved by downloading or generating type definitions for all upstream dependencies of every downstream dependency; this, however, would reduce the performance of the approach.
	\item In general, the static type analyzer of TypeScript has its limitations.
		For instance, the type analysis will have a limited recall for certain control flow patterns or metaprogramming constructs such as meta-circular evaluation using the the \code{eval()} function or dynamic function binding using \code{Function.bind()} etc.
\end{itemize}

@@

To answer \cref{q2,q3} without our tool, users need to scan each found dependency separately to count or read all usage samples.
If they are intending to analyze the usage of a member with an unambiguous name, they can perform an expeditious string search.
This is possible using Sourcegraph if the dependency is indexed there; otherwise, users will need to download the repository and search it locally.
If the member of interest has an ambiguous name, users will need to download the project and view it in an IDE such as VS Code that supports reference search.
Alternatively, they can try the precise code intelligence mode on Sourcegraph (see \cref{sec:related_work/usage_samples}); however, this feature is available for a small number of repositories only.

@@

The spatial distance varies for the different views of the tool:
For the dependency browser and the usage browser, a separate toolbar has to be opened, which indicates a higher spatial distance; still, the artifacts are available in the same IDE.
Also, the dependency artifacts have no strong relationship to any existing IDE artifacts, so any closer integration into an existing view would increase coupling between independent domains.
The code annotations are disaplyed close to the original source code with a small spatial distance.

>>>>

\begin{enumerate}[label=RQ\arabic*]
	\item \label[requirement]{resqu1} What is the quality and quantity of the proposed methods for dependency collection?
	\item \label[requirement]{resqu2} What is the quality and quantity of the proposed method for mining usage samples?
	\item \label[requirement]{resqu3} How well is the proposed tool applicable?
\end{enumerate}

@@

To assess the quality and quantity of dependency collection, we analyze it with respect to extent, precision, recall, and performance.
\Cref{tab:evaluation/resqu1/quantities} shows the number of dependencies that we collected to build an experimental dataset from each data source following the proposed methods for a set of arbitrarily selected npm packages.
For each package, we have annotated a subset of the collected dependencies (max. 20 dependencies per package) to identify and classify false positive hits.

\begin{table}
	\caption{Quantity and false-positive rates (FPR) of downstream dependencies found by both methods for selected packages.}
	\label{tab:evaluation/resqu1/quantities}

@@

Regarding the total number of dependencies collected by each method, no clear trend in favor of any method can be ascertained for small or medium packages (having less than \num{10,000} stars on GitHub).
For larger packages (having at least \num{10,000} stars), however, the number of dependencies collected from the npm registry stagnates near to 400 hits; beyond this limit, the npm registry will return internal server errors only.
For Sourcegraph, we have not hit any limitations so far; currently, they do not provide any official documentation for the exact rate limits.

As the disjunct proportion of results from both methods is very small (on average about \SI{6}{\percent}), we consider a combination of both data sources useful for maximizing the extent and diversity of the gained dataset.

@@

In some situations, this may happen if the package is a plugin for another package or if it is invoked as a CLI from a build script, but in the majority of repositories, developers trivially appear to have specified the upstream dependency by accident (e.g., while copying a \code{package.json} file over from another package), or to have forgotten to remove the upstream dependency after switching away from using the package.
Rarer causes include repositories that only add type definitions for TypeScript to a package but do not actually import it, as well as, especially for repositories found on Sourcegraph, other packages required by a repository that declares a peer dependency on the target package which needs to be fulfilled by the depending repository.

@@

Besides the precision of collected dependencies, their recall is also of interest.
As the dependencies are collected in excerpts from two very large datasets, a quantitative analysis of false-negative hits based on manual annotation of these datasets would be too expensive for this paper.
Nevertheless, there are some causes that will prevent a dependency from being found by our methods:

\begin{itemize}
	\item Packages without a proper dependency manifest cannot be detected by our methods which are based on exactly these metadata.
	\item On npm, only published packages will be found, creating a bias towards generic and professional software.
	\item On both platforms, results will be sorted based on intransparent criteria (which, as we speculate, include the number of direct dependents on npm and the recent update frequency on Sourcegraph, resp.).
		As we only fetch the first dependencies from both data sources, this is a likely source of further biases.
		These biases could be fought by always fetching all dependencies, but this would drastically reduce the overall performance.
\end{itemize}

@@

To assess the quality and quantity of usage mining, we examine the precision, recall, and performance of this processing step.
Due to the complexity of a detailed annotation process, we only consider the existence of found usage samples on the dependency level.
The absolute quantity of found usage samples largely varies for different packages and dependencies based on the semantic extent of the package and the coupling between package and dependency.

@@

Under laboratory conditions, false-positive usage samples cannot be emitted by our method if we assume the correctness of the TypeScript compiler.
There are only two theoretical exceptions to this invariant:
\begin{enumerate*}[label=(\roman*)]
	\item In case of a name collision between two packages, invalid or false positives usage samples may be emitted.
		However, the npm infrastructure attempts to rule out these collisions by using the package name as a unique ID for every published package.
	\item If dependency developers deliberately interfere with the TypeScript compiler (for instance, by suppressing type errors using a \verb|@ts-ignore| comment\footnote{\url{https://www.typescriptlang.org/docs/handbook/release-notes/typescript-2-6.html\#suppress-errors-in-ts-files-using--ts-ignore-comments}}), both false positive and false negative matches are possible.
\end{enumerate*}

@@

To evaluate the recall of our method, we have refined the experimental dataset collected in \cref{sec:evaluation/resqu1} and removed all false-positive dependencies.
Within this cleansed dataset, the overall rate of dependencies for that our method is unable to find at least one usage sample accounts for \SI{47.9}{\percent}.
To explain these false negatives, we have identified a number of systematic causes:

\begin{itemize}
	\item Many projects on OSS platforms have complex build configurations that include additional transpilation or code generation steps before a final version of the source code is reached that is valid to the node.js interpreter or the TypeScript compiler.
		Unless we add explicit support for such build configurations, type analysis and usage mining for these projects will fail.
	\item In some situations, additional type definitions are required to perform a complete type inference of a dependency.
		This applies to every parametrized callback from a package for that type definitions are not available.
		This limitation could be resolved by downloading or generating type definitions for all upstream dependencies of every downstream dependency; this, however, would reduce the performance of the approach.
	\item In general, the static type analyzer of TypeScript has its limitations.
		For instance, the type analysis will have a limited recall for certain control flow patterns or metaprogramming constructs such as meta-circular evaluation using the \code{eval()} function, dynamic function binding using \code{Function.bind()}, etc.
\end{itemize}

@@

To answer \cref{q2,q3} without our tool, users need to scan each found dependency separately to count or read all usage samples.
If they are intending to analyze the usage of a member with an unambiguous name, they can perform an expeditious string search.
This is possible using Sourcegraph or another search engine if the dependency is indexed there; otherwise, users will need to download the repository and search it locally.
If the member of interest has an ambiguous name, users will need to download the project and view it in an IDE such as VS Code that supports reference search.
Alternatively, they can try the precise code intelligence mode on Sourcegraph (see \cref{sec:related_work/usage_samples}); however, this feature is available for a small number of repositories only.

@@

The spatial distance varies for the different views of the tool:
For the dependency browser and the usage browser, a separate toolbar has to be opened, which indicates a higher spatial distance; still, the artifacts are available in the same IDE.
Also, the dependency artifacts have no strong relationship to any existing IDE artifacts, so any closer integration into an existing view would increase coupling between independent domains.
The code annotations are displayed close to the original source code with a small spatial distance.

### 2021-10-11 docs/paper/sections/6_conclusion.tex
Minor rephrasings and fixed throughall the paper

Our analysis suggests that the collected dependencies and the extracted usage samples have a viable quality in terms of precision and extent, and that the resource requirements of our implementation are sufficiently low for collecting all data on the local machine of a package developer.
However, we have identified a number of causes for false negatives over the whole mining process that are predestined to biasing the output usage samples, including the popularity and recency of dependencies but also the proper declaration of packages and their renunciation of complex toolchains and metaprogramming patterns.
Our analysis, however, is not yet fully supported by data and would require further quantitative evaluation to assess the exact precision and recall of our approach for mining usage samples as well as the efficacy of our tool.
As for the first question, manual annotation of source code repositories with usage samples will be required to compare them to the outputs of our implementation.
Close insights for the second question could be gained from a user study that compares the efficiency of package developers asked to solve a set of downstream dependency-related tasks with or without the help of our tool.

@@

In a wider perspective, we envision many other purposes for analyzing the collected downstream dependency data.
By deriving usage patterns from the samples using an approach such \cite{zhong2009mapo}, the structure and classification of raw data could be refined.
The derived patterns could also be exploited as suggestions for convenience protocols to package developers.
There are many ways to turn dependency data into metrics that either could be used for classifying downstream dependencies (e.g., according to their coupling to the target package, LCOM), or that could be attributed to ranges of packages for comparing them with regard to their dependency-based cohesion or their spreading in an ecosystem.

By replacing the static usage analysis with a dynamic approach, the detect usage samples could be enriched with valuable runtime data describing involved parameters, the fine-grained code coverage of package members, or the invocation context of members.
Returning to the abstract problem that we have formulated in the introduction -- how can we improve developer knowledge about their packages' usage? --, further data sources next to source code files emerge that could be mined for package references, too.
For instance, we are looking forward to mining the change history of projects for package-related changes (potentially indicating breaking changes), searching conversation platforms such as public issue trackers or Q\&A forums for mentions of package members and analyzing their sentiments (potentially indicating confusing behavior or bad documentation), or even scanning continuous integration logs for failure stack traces (potentially revealing bugs in the package of interest).

>>>>

Our analysis suggests that the collected dependencies and the extracted usage samples have a viable quality in terms of precision and extent, and that the resource requirements of our implementation are sufficiently low for collecting all data on the local machine of a package developer.
Nevertheless, we have identified a number of causes for false negatives over the whole mining process that are predestined to biasing the output usage samples, including the popularity and recency of dependencies but also the proper declaration of packages and their renunciation of complex toolchains and metaprogramming patterns.
Our analysis, however, is not yet fully supported by data and would require further quantitative evaluation to assess the exact precision and recall of our approach for mining usage samples as well as the efficacy of our tool.
As for the first question, manual annotation of source code repositories with usage samples will be required to compare them to the outputs of our implementation.
Close insights for the second question could be gained from a user study that compares the efficiency of package developers asked to solve a set of downstream dependency-related tasks with or without the help of our tool.

@@

In a wider perspective, we envision many other purposes for analyzing the collected downstream dependency data.
By deriving usage patterns from the samples using an approach such as \cite{zhong2009mapo}, the structure and classification of raw data could be refined.
The derived patterns could also be exploited as suggestions for convenience protocols to package developers.
There are many ways to turn dependency data into metrics that either could be used for classifying downstream dependencies (e.g., according to their coupling to the target package as per LCOM \citep{chidamber1994metrics}), or that could be attributed to ranges of packages for comparing them with regard to their dependency-based cohesion or their spreading in an ecosystem.

By replacing the static usage analysis with a dynamic approach, the detected usage samples could be enriched with valuable runtime data describing involved parameters, the fine-grained code coverage of package members, or the invocation context of members.

Returning to the abstract problem that we have formulated in the introduction -- how can we improve developer knowledge about their packages' usage? --, further data sources next to source code files emerge that could be mined for package references, too.
For instance, we are looking forward to mining the change history of projects for package-related changes (potentially indicating breaking changes), searching conversation platforms such as public issue trackers or Q\&A forums for mentions of package members and analyzing their sentiments (potentially indicating confusing behavior or bad documentation), or even scanning continuous integration logs for failure stack traces (potentially revealing bugs in the package of interest).
